---
title: "DS6306_Case_Study_2"
author: "Triston Hudgins"
date: "11/26/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(readxl)
library(dplyr)
library(tidyr)
library(tidyverse)
library(knitr)
library(ggplot2)
library(GGally)
library(ggthemes)
library(png)
library(ggimage)
library(naniar)
library(plotly)
library(class)
library(caret)
library(e1071)

```


#Pull Data

```{r}

#Import Datasets
CaseStudy2 <- read_csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2-data.csv")

AttritionTest <- read_csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2CompSet%20No%20Attrition.csv")

IncomeTest <- read_excel("CaseStudy2CompSet No Salary.xlsx")

```

<br>

<br>

#Missing Values Check and initial plotting group
```{r}

#check missing values
gg_miss_var(CaseStudy2)

#Initial plotting group
CaseStudy2 %>%
  select(Attrition, MonthlyIncome, JobInvolvement, DistanceFromHome, YearsInCurrentRole, WorkLifeBalance, OverTime) %>%
  ggpairs(aes(color=Attrition))

```

<br>

<br>

#Attrtion by Department
<br>
##As each department generally has different challenges, we have elected to observe attrtion by departments and job roles

```{r}

#Analyze overall attrtion
Attritiondf <- CaseStudy2 %>%
                  filter(Attrition == "Yes")

#Attrition Counts by Department
Attritiondf %>%
  ggplot(aes(x=Attrition, fill = Attrition)) +
  geom_bar(alpha = 0.6) +
  scale_fill_manual(values=c('#E20A17')) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank(),
        plot.title = element_text(size = 15, face = "bold"),
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold")) +
  labs(title = "Bar Plot of Attrition Count by Department", y = "Count") +
  facet_grid(~Department)
#Sales and Research and Development have the highest attrition count 

#Attrition Counts by Job Role
Attritiondf %>%
  ggplot(aes(x=Attrition, fill = Attrition)) +
  geom_bar(alpha = 0.6) +
  scale_fill_manual(values=c('#E20A17')) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank(),
        plot.title = element_text(size = 15, face = "bold"),
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold")) +
  labs(title = "Bar Plot of Attrition Count by Job Role", y = "Count") +
  facet_grid(~JobRole)

#Counts and percentages of attrition by Department
Department <- CaseStudy2 %>%
                group_by(Department) %>%
                count(Department, Attrition) %>%
                mutate(Percent = (n / sum(n))*100) %>%
                ungroup()

kable(Department)

#Percent Atrition by Department
PercentAttritionDP <- CaseStudy2 %>%
  group_by(Department) %>%
  count(Department, Attrition) %>%
  mutate(Percent = (n / sum(n))*100) %>%
  ungroup()

#Attrition by Department Percentage
PercentAttritionDP %>%
  filter(Attrition == "Yes") %>%
  group_by(Department) %>%
  ggplot(aes(x=Attrition, y=Percent, fill = Attrition)) +
  geom_bar(stat = "identity", alpha = 0.6) +
  scale_fill_manual(values=c('#E20A17')) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank(),
        plot.title = element_text(size = 15, face = "bold"),
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold")) +
  labs(title = "Bar Plot of Attrition Percentage by Job Role", y = "Percent") +
  facet_grid(~Department)

#Percent Atrition by Job Role
PercentAttritionJR <- CaseStudy2 %>%
  group_by(JobRole) %>%
  count(JobRole, Attrition) %>%
  mutate(Percent = (n / sum(n))*100) %>%
  ungroup()

#Attrition by Job Role Percentage
PercentAttritionJR %>%
  filter(Attrition == "Yes") %>%
  group_by(JobRole) %>%
  ggplot(aes(x=Attrition, y=Percent, fill = Attrition)) +
  geom_bar(stat = "identity", alpha = 0.6) +
  scale_fill_manual(values=c('#E20A17')) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank(),
        plot.title = element_text(size = 15, face = "bold"),
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold")) +
  labs(title = "Bar Plot of Attrition Percentage by Job Role", y = "Percent") +
  facet_grid(~JobRole)

#Job Role Percent Attrition
JobRoleCount <- CaseStudy2 %>%
                  group_by(JobRole) %>%
                  count(JobRole, Attrition) %>%
                  mutate(Percent = (n / sum(n))*100) %>%
                  ungroup()

kable(JobRoleCount)

```

<br>

#Monthly Income By Job Role

<br>

```{r}

#BoxPlot of Attrition and Monthly Income
CaseStudy2 %>%
  ggplot(aes(x=MonthlyIncome, fill = Attrition)) + 
  geom_boxplot(alpha = 0.6) + 
  coord_flip() +
  scale_fill_manual(values=c('#EBC301','#E20A17')) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "bottom",
        plot.title = element_text(size = 15, face = "bold"),
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold"),
        rect = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  labs(title = "Box Plot of Monthly Income by Job Role Attrition") +
  facet_grid(~JobRole)
#Largest disparity is in HR
#HR, Lab Tech, and Sales Rep all have lower median monthly income among the attrition group


#Filter HR, Lab Tech, and Sales Rep
target <- c("Sales Representative", "Laboratory Technician", "Human Resources")

CaseStudy2 %>%
  filter(JobRole %in% target) %>%
  ggplot(aes(x=MonthlyIncome, fill = Attrition)) + 
  geom_boxplot(alpha = 0.6) + 
  coord_flip() +
  scale_fill_manual(values=c('#EBC301','#E20A17')) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "bottom",
        plot.title = element_text(size = 15, face = "bold"),
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold"),
        rect = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  labs(title = "Box Plot of Monthly Income by Job Role Attrition") +
  facet_grid(~JobRole)

#Show Job Role and Attrition Stats
CaseStudy2 %>%
  filter(JobRole %in% target) %>%
  group_by(JobRole, Attrition) %>%
  summarise(Mean=mean(MonthlyIncome), Max=max(MonthlyIncome), Min=min(MonthlyIncome),
            Median=median(MonthlyIncome), Std=sd(MonthlyIncome))

#T Test for difference in mean monthly income in attrition group
TTest <- CaseStudy2 %>%
            filter(JobRole %in% target)

t.test(CaseStudy2$MonthlyIncome~CaseStudy2$Attrition)

t.test(TTest$MonthlyIncome~TTest$Attrition)



```

<br>

#Overtime By Job Role

<br>

```{r}

#Percent Attrition by Overtime and Job Role
PercentAttrition <- CaseStudy2 %>%
  group_by(JobRole) %>%
  count(JobRole, OverTime, Attrition) %>%
  mutate(Percent = (n / sum(n))*100) %>%
  ungroup()

#Box Plot of Monthly Income by Overtime Requirement
CaseStudy2 %>%
  group_by(OverTime) %>%
  filter(JobRole %in% target) %>%
  ggplot(aes(x=MonthlyIncome, y = OverTime, fill = Attrition)) + 
  geom_boxplot(alpha = 0.6) + 
  coord_flip() +
  scale_fill_manual(values=c('#EBC301','#E20A17')) +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 15, face = "bold"),
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold"),
        rect = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  labs(title = "Box Plot of Monthly Income by Overtime Requirement") +
  facet_grid(~JobRole)

#Show Overtime and Attrition Stats
CaseStudy2 %>%
  filter(JobRole %in% target) %>%
  filter(OverTime == "Yes") %>%
  group_by(JobRole, OverTime, Attrition) %>%
  summarise(Mean=mean(MonthlyIncome), Max=max(MonthlyIncome), Min=min(MonthlyIncome),
            Median=median(MonthlyIncome), Std=sd(MonthlyIncome))

```

<br>

#Job Involvement By Job Role

<br>

```{r}


#Percent Attrition by Job Involvement and Job Role
PercentAttrition <- CaseStudy2 %>%
  group_by(JobRole) %>%
  count(JobRole, JobInvolvement, Attrition) %>%
  mutate(Percent = (n / sum(n))*100) %>%
  ungroup()

kable(PercentAttrition)

PercentAttrition %>%
  group_by(JobRole) %>%
  ggplot(aes(x=JobInvolvement, y=Percent, fill = Attrition)) +
  geom_bar(stat = "identity", alpha = 0.6) +
  scale_fill_manual(values=c('#EBC301','#E20A17')) +
  theme(plot.title = element_text(size = 15, face = "bold"),
        legend.position = "bottom",
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold")) +
  labs(title = "Bar Plot of Attrition Percentage by Job Role and Job Involvement", y = "Percent") +
  facet_grid(~JobRole)

#Box Plot of Monthly Income by Job Involvement
CaseStudy2 %>%
  group_by(JobInvolvement) %>%
  filter(JobRole %in% target) %>%
  ggplot(aes(x=MonthlyIncome, y = as.factor(JobInvolvement), fill = Attrition)) + 
  geom_boxplot(alpha = 0.6) + 
  coord_flip() +
  scale_fill_manual(values=c('#EBC301','#E20A17')) +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 15, face = "bold"),
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold"),
        rect = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  labs(title = "Box Plot of Monthly Income by Job Involvement", y = "Job Involvement Level") +
  facet_grid(~JobRole)

```

<br>

#KNN Modeling to Predict Attrition

<br>

```{r}

####Attrition Prediction Analysis####

KNNtarget <- c( "ID", "BusinessTravel", "Over18", "StandardHours", "Department", "EducationField", "Gender", "JobRole", "MaritalStatus", "OverTime")

Train = CaseStudy2[, !names(CaseStudy2) %in% KNNtarget]
Test = AttritionTest[, !names(AttritionTest) %in% KNNtarget]

data.frame(Train)
data.frame(Test)

Train[sapply(Train, is.character)] <- lapply(Train[sapply(Train, is.character)], as.factor)

str(Train)
str(Test)

# K = 3
classifications = knn(Train[, names(Train) != "Attrition"], Test, Train$Attrition, prob = TRUE, k = 3)
KNNTable = table(Test$EmployeeNumber,classifications)

#Change to dataframe
KNNTable <- as.data.frame.matrix(KNNTable) 

KNNTable <- setNames(cbind(rownames(KNNTable), KNNTable, row.names = NULL), c("EmployeeNumber", "No", "Yes"))

#Set up for Merge
KNNTable <- KNNTable %>%
              select(EmployeeNumber, Yes) %>%
              rename(Attrition = Yes) %>%
              mutate(Attrition = replace(Attrition, Attrition == 1, "Yes")) %>%
              mutate(Attrition = replace(Attrition, Attrition == 0, "No"))

AttritionTest <- merge(AttritionTest, KNNTable, by ="EmployeeNumber", all=TRUE)

#write.csv(AttritionTest,"/Users/thud1/OneDrive/Documents/DS6306 Doing Data Science/Case Study 2//Case2PredictionsHudgins-Attrition.csv", row.names = FALSE)

#Prediction Attrition Counts by Job Role
AttritionTest %>%
  ggplot(aes(x=Attrition, fill = Attrition)) +
  geom_bar(alpha = 0.6) +
  scale_fill_manual(values=c('#EBC301','#E20A17')) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank(),
        plot.title = element_text(size = 15, face = "bold"),
        legend.position = "bottom",
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold")) +
  labs(title = "Bar Plot of Predicted Attrition Count by Job Role", y = "Count") +
  facet_grid(~JobRole)

```

<br>

#Multiple Linear Regression to Predict MOnthly Income

<br>

```{r}

#Set up model
fit = lm(MonthlyIncome~JobLevel + TotalWorkingYears, CaseStudy2)
summary(fit)

IncomeTestlm <- IncomeTest %>%
                select(JobLevel, TotalWorkingYears)

#Get predictions
predictions <- fit %>% 
                predict(data.frame(IncomeTestlm))

LMTable = as_tibble(predictions)

LMTable <- LMTable %>%
              rename(MonthlyIncome = value)

IncomeTestFull <- merge(IncomeTest, LMTable, by = 0)

IncomeTestFull <- IncomeTestFull[order(IncomeTestFull$ID),]

#write.csv(IncomeTestFull,"/Users/thud1/OneDrive/Documents/DS6306 Doing Data Science/Case Study 2//Case2PredictionsHudgins-Salary.csv", row.names = FALSE)

#BoxPlot of Attrition and Predicted Monthly Income
IncomeTestFull %>%
  ggplot(aes(x=MonthlyIncome, fill = Attrition)) + 
  geom_boxplot(alpha = 0.6) + 
  coord_flip() +
  scale_fill_manual(values=c('#EBC301','#E20A17')) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "bottom",
        plot.title = element_text(size = 15, face = "bold"),
        axis.title = element_text(face = "bold"),
        strip.text.x = element_text(face = "bold"),
        rect = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  labs(title = "Box Plot of Predicted Monthly Income by Job Role Attrition") +
  facet_grid(~JobRole)

```

##### Conclusion
We have determined that the leading influences on attrition are monthly income disparities, over time requirements, and job involvement.
